### 2017-04-24 main.tex

\documentclass[10pt,conference,compsoc]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[nocompress]{cite}
\usepackage{caption}
\usepackage{subcaption}

%\newcommand{\per}[1]{}
%\newcommand{\anna}[1]{}
%\newcommand{\kiran}[1]{}

% Title Page
\title{Probe or Wait: Handling tail losses using Multipath TCP}

\author{\IEEEauthorblockN{Kiran~Yedugundla, Per~Hurtig, Anna~Brunstrom}
\IEEEauthorblockA{Dept. of Computer Science, Karlstad University, Karlstad, Sweden}}

\begin{document}
\maketitle

\begin{abstract}
Packet losses are known to affect the performance of latency sensitive applications in the Internet such as media streaming and gaming. Transport protocols recover from packet loss in order to provide reliable end to end communication and improving the quality of user experience. The efficiency of loss recovery greatly influences the completion time of flows. In this paper we focus on the state of the art loss recovery mechanisms for TCP and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms. Based on the observations, we propose an enhancement to the tail loss recovery in Multipath TCP to improve the loss recovery time. Our experiment results show consistent end to end latency performance improvement in considered scenarios. 
\end{abstract}

\section{Introduction}

Multipath TCP~(MPTCP) is an experimental standard proposed as an extension to TCP~\cite{rfc6824}. It allows devices with multiple interfaces to transmit data simultaneously on both interfaces, thereby improving the throughput of end-to-end connections. Using MPTCP, a connection can have multiple TCP subflows using different interfaces on different routes. Each TCP subflow imparts the functional behavior of a standard TCP flow.

TCP has two mechanisms for detecting and recovering from packet losses: Fast Retransmit (FR) and Retransmission Timeout~(RTO)~\cite{Flach:2013}. Fast retransmit is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as a loss indication. In case of RTO, a sender waits for the packet loss to trigger a timeout before retransmitting that packet. During timeout, the sender cannot send any packets, but with FR the out-of-order packets trigger duplicate acknowledgements. The sender can start retransmission of missing packets without waiting for timeout. For flows where there are sufficiently many packets to be transferred, FR is quicker than RTO in detecting and recovering packet losses. Fast retransmit coupled with fast recovery can send data outstanding when there is enough data to send in long flows. On the other hand, shorter flows more frequently depend on RTO for packet loss detection and recovery, thus incurring more latency. A single packet loss in a short flow may take many RTTs to detect and recover. This scenario is also applicable to the packets at the end of the flow (or tail), in a long flow. Although, RTO is reliable, even in tail loss~(losses at the end of transmission) recovery, it is not efficient in terms of recovery completion time.

TCP Loss Probe (TLP)~\cite{Flach:2013} is a mechanism to detect and recover from tail losses by employing a loss probe packet for a timeout shorter than the RTO. Probe timeout enables short flows to recover packet loss at the tail, by triggering FR. It assumes that other algorithms such as Early Retransmit~(ER)\cite{rfc5827} and Forward Acknowledgement~(FACK)~\cite{FACK} threshold based recovery are enabled. TLP is available in Linux kernel and enabled by default. Further, evaluations in~\cite{Rajiullah:2015} showed that TLP provides significant reductions in latency for short flows with one to n-degree tail losses.

 MPTCP connections start with one TCP flow to send data over, and may add more flows to the connection if configured to. MPTCP uses a connection level congestion window as well as subflow level congestion windows for each subflow. Data is split across the flows and each data packet is scheduled on one of the subflows based on minimum RTT. Packet losses in each subflow are detected and recovered in a similar fashion as that of TCP. Loss recovery can be conducted at different levels: the MPTCP connection level (meta level) and at subflow level. If the recovery is handled at meta level, a lost packet may be rescheduled and retransmitted on the available subflow with the lowest RTT. If recovery is handled at the subflow level, the packet must be retransmitted on the same subflow. An important part of loss recovery process at subflow level is to follow the TCP semantics and retransmit lost packets on the initial path of transfer. However, the MPTCP standard~\cite{rfc6824} does not restrict on, how the loss recovery should happen in the implementation and which subflow should retransmit the lost packets.

In this paper, we analyze the retransmission behavior of MPTCP in selected tail loss cases to identify the scope for improvement. We propose a change in the retransmission strategy of MPTCP to improve the latency performance. We provide related research on MPTCP retransmission in section~\ref{relwork}, and describe the experimental setup in section~\ref{exsetup}. We discuss observations from the experiments on state of the art retransmission behavior of TCP and MPTCP in section~\ref{disc}. Based on the observations, we propose an improvement to the current Linux implementation of TLP for MPTCP in section~\ref{impr}. In section~\ref{eval}, we provide an evaluation of the proposed TLP for MPTCP mechanism and conclude the paper in section~\ref{conc}.

%\kern-1em
\section{Background and Related Work}\label{relwork}

Retransmission schemes are of significant importance in achieving low flow completion time. Several IETF RFCs such as SACK based recovery~\cite{rfc6675}, Limited Transmit~\cite{rfc3042} and Early Retransmission~\cite{rfc5827} have led to improved retransmission strategies for TCP. TLP is part of the various deployable loss recovery mechanisms proposed in~\cite{Flach:2013}, suggesting that current Linux TCP implementation provides room for improvement in several packet loss scenarios.

In a TCP flow, sequence numbers are used to identify lost packets and retransmit them. MPTCP uses two levels of sequence numbers to support efficient data transfer, namely data sequence and TCP sequence. Data sequence numbers are for the end-to-end data transfer and TCP sequence numbers are for individual flow data sequence. There is an association between data sequence numbers and TCP sequence numbers within a subflow.  A loss occurring in an individual TCP flow corresponds to a loss in end-to-end data. Multipath TCP, as an extension of TCP, uses the basics of TCP retransmission. However, the interaction with two levels makes the problem of retransmission more challenging than in TCP. The dual sequence numbering enables MPTCP to respond to loss of packets by retransmitting the lost packets on an alternate path.

The Linux implementation of MPTCP uses a set of retransmission heuristics to handle retransmissions. The data outstanding on a timed out subflow should be rescheduled for transmission on a different subflow using timeout as the indicator. Fast retransmit on a subflow will not trigger retransmission on another subflow. The target devices of MPTCP contain a degree of asymmetry in their interface characteristics such as 3G or 4G vs WLAN, leading to flows with different delay characteristics. Significant delay differences in available paths can cause receive buffer limitations, as the receiver has to wait for the data sent on a slow path, while buffering the data on the fast path. In such cases, opportunistic retransmission~\cite{Costin} improves the latency by using fast path to retransmit the data sent on an another path. Authors in~\cite{fuso}, try to exploit the path diversity by quickly retransmitting on the fastest paths. Such quick retransmission comes with a cost of redundant packets as each individual TCP flow should retransmit always as well to follow TCP semantics. The MPTCP standard~\cite{rfc6824} suggests a more conservative approach. MPTCP is designed to exploit the multiple interfaces and often one or more of these could be wireless with completely different characteristics than that of wired interfaces. Authors in~\cite{Shin} argue that the calculation of RTO for MPTCP flows should include the interface characteristics to improve loss recovery time.

%\subsection{Scope}\label{scope}

This paper analyzes the performance of state of the art loss recovery mechanism of MPTCP and presents a case for improvement in the tail loss scenario. 
For understanding the loss recovery and retransmission policies in MPTCP Linux implementation version 0.91, we consider several specific cases of tail losses.  
We use the standard settings of the Linux kernel.
The Linux implementation of TCP is considered as the baseline TCP retransmission for implementation flexibility and availability of TLP support.

%\kern-1em
\section{Experimental Setup}\label{exsetup}

Experiments use the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}. The client node connect to two wireless interfaces 3G or 4G and WLAN and the server node connect to a wired router.
CORE uses the Linux networking stack in each node of the topology by using a lightweight virtualization technique known as Linux network namespaces.
Characteristics of the connection setup and assumptions about the parameters are provided in Table~\ref{tab1}.
%The parameters are chosen to be close to real values of bandwidth and delay of respective technologies.
There is no attempt in this study to focus on the effect of link detailed characteristics in retransmission performance.

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.45\textwidth]{images/fortest.pdf}
\caption{Topology used for Emulation}\label{fig1}
\end{center}
\end{figure}
\begin{center}

\begin{table}
\resizebox{0.45\textwidth}{!}{\begin{minipage}{0.45\textwidth}
\small
\begin{center}
\begin{tabular}{|c|cccccccccc|}
      \hline

      \multicolumn{1}{c}{} & & \\[\dimexpr-\normalbaselineskip-\arrayrulewidth]
      \textbf{Burst Size} & \multicolumn{10}{c|}{80 Packets} \\
      \hline
      \textbf{Separation Time} & \multicolumn{10}{c|}{2s} \\
      \hline

      \textbf{Oneway Delay} & \multicolumn{10}{c|}{20\,ms-120\,ms}  \\
      \hline 	
      \textbf{Bandwidth} & \multicolumn{10}{c|}{54Mbps}  \\
      \hline
      \textbf{Loss Model} & \multicolumn{10}{c|}{Deterministic}\\
      \hline

\end{tabular}
\caption{Emulation parameters}\label{tab1}
\end{center}
\end{minipage}}
\end{table}
\end{center}

%\kern-1em
\subsection{Testing retransmission with deterministic loss patterns}
In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
Losses are generated by associating netem with corresponding interfaces and dropping select packets. This process is simplified by using the KAUNetem tool~\cite{Garcia2016}, which allows experimenter to e.g., drop packets based on their positions in a flow. A MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between the client and the server. So one has to wait both in time and packets for the second subflow establishment. We send two bursts of 80 packets each and drop tail packets in the second burst at one of the interfaces. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that subflow. We calculate the burst completion time for the second burst that has specified tail losses. The objective of this experiment is to evaluate retransmission mechanisms with tail loss (single packet loss and multiple tail loss) and temporary path loss. Three test cases are considered for experiments: single packet tail loss and two packet tail loss for tail loss evaluation and single packet along with probe~(in TLP) or retransmission~(in RTO) loss for path loss evaluation.

In Linux, TCP supports multiple retransmission mechanisms that can be controlled by a system setting. The default setting in Linux enables both
ER and TLP, which may not be the case with other operating systems. In this paper, we use TLP to denote the default Linux setting and RTO to denote the use of retransmission timeout only.

The setup has server and client with Linux supporting MPTCP running on them. The Client has two interfaces with one way delay on each interface ranging from 20\,ms to 120\,ms. The intention is to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in the link delays. We consider 5 scenarios with delay pairings 20\,ms-20\,ms, 20\,ms-30\,ms, 20\,ms-120\,ms, 30\,ms-20\,ms, 120\,ms-20\,ms to understand the effect of delay difference in the performance as shown in Table~\ref{tab1}.

\kern-1em
\section{Observations and Discussion}\label{disc}
This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, two packet tail loss and retransmission or probe loss.
\kern-5em
\subsection{TCP}
Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
starts with the results using TCP and compare it with that of MPTCP. Figure~\ref{tcpf} represent the comparison of TCP 
burst completion times. Path asymmetry is irrelevant in this case as TCP uses a single path.

In the case of single packet tail loss, there is no difference in burst completion times of RTO and TLP settings. The retransmission
timeout and the probe timeout are the same as there is only one outstanding packet. This is a special case in the TLP
specification that adds 200\,ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts
for the delayed ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for PTO.

Hence retransmission timeout is the mode of loss recovery in both RTO and TLP. This can be avoided by tuning
the ACKs as discussed in~\cite{Rajiullah:2015}. In the case of two packet tail loss, TLP performs better than RTO as it employs a combination of TLP and ER to recover from losses. 
If the probe or retransmission is lost, then TLP still performs better than RTO as it waits for one probe timeout and one RTO instead of two RTOs, thus avoiding the exponential backoff of the RTO.

\begin{figure*}[!tbp]
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T1P.pdf}
	\caption{Single packet}\label{t1p}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/T2P.pdf}
	\caption{Two packets }\label{t2p}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/T1PP.pdf}
	\caption{Single packet and probe loss }\label{t1pp}
 \end{subfigure}
 \caption{Tail loss scenarios using TCP}\label{tcpf}
\end{figure*}
%\kern-1em
\subsection{MPTCP}

The expected retransmission behavior of MPTCP for RTO and Linux default settings is shown 
in Figure.~\ref{mptiming}. The actual pattern might differ in cases, where the RTT is larger than 200\,ms. 
\renewcommand{\thesubfigure}{\roman{subfigure}}
\begin{figure*}[!tbp]
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing1P.pdf}
	\caption{Single packet}\label{timing1P}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth} 
	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing2P.pdf}
	\caption{Two packets }\label{timing2P}
 \end{subfigure} 
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
  	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing1PP.pdf}
	\caption{Single packet and probe loss}\label{timing1PP}
 \end{subfigure}
 \caption{Timing diagrams of MPTCP behavior on tail losses with a) RTO b) TLP }\label{mptiming}	
\end{figure*}

MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit
the data until its congestion window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance
is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO and TLP for 
cases when both paths are symmetric or the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens in the second path with lowest RTT following the behavior depicted
in Figure~\ref{timing1P}.

\renewcommand{\thesubfigure}{\alph{subfigure}}
\begin{figure*}[!tbp]
 \begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/1P.pdf}
\caption{Single packet}\label{1p}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth,natwidth=578.16,natheight=433.62]{plots/2P.pdf}
\caption{Two packets}\label{2p}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/1PP.pdf}
\caption{Single packet and probe loss}\label{1pp}
 \end{subfigure}
 \caption{Tail loss scenarios using MPTCP}
\end{figure*}

In the case of single packet tail loss, the last outstanding packet is sent as probe and in case of n-degree tail loss,
the latest packet is sent as probe to trigger fast retransmit of the packets in between. In order to observe
the performance of n-degree tail loss, we consider dropping the last two packets. 
The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the other 
two settings in all scenarios except when the packets are lost on a high delay interface. In this scenario, the MPTCP retransmission follows the 
behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of the last packet on the alternate path is 
faster than the probe on the primary path.

To further study the performance of TLP in the case when there is longer path interruption on one path, we tried to drop the probe packet along with the last packet. In this scenario, the RTO
setting results in much lower burst completion times. Comparing with the performance of TLP in TCP from Figure~\ref{t1pp}, where we saw that use of TLP improved the burst completion time, TLP in MPTCP did not improve the performance. The reason lies in the way the retransmission and loss probe transmission is carried out in MPTCP. Retransmission occurs after an RTO time on both paths.
Even if the retransmission fails on one path, the client receives the packet on the other path as shown in Figure~\ref{timing1PP}. For TLP, in
current implementation, the loss probe is sent on the actual path of the first transmission but not on both paths. This incomplete impartation of
TLP from TCP to MPTCP led to the increase in burst completion time by an RTO for MPTCP.

%\begin{table}[!ht]
%\centering
%\caption{Tail loss scenarios with tcp.early.retrans = 3 default}
%\label{ret3}
%\begin{tabular}{|l|l|l|l|l|}
%\hline
% testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
%Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
%Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
%Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                &  \\ \hline
%\end{tabular}
%\end{table}

%\begin{table}[!ht]
%\centering
%\caption{Tail loss scenarios with tcp.early.retrans = 0 ER disabled}
%\label{ret0}
%\begin{tabular}{|l|l|l|l|l|}
%\hline
% testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
%Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
%Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
%Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
%Tail drop 4 & 20ms - 120ms &  			&		&  \\ \hline
%Tail drop 5 & 120ms - 20ms &  			& 		& \\ \hline 
%\end{tabular}
%\end{table}

%\begin{table}[!ht]
%\centering
%\caption{Tail loss scenarios with tcp.early.retrans = 4 TLP disabled}
%\label{ret4}
%\begin{tabular}{|l|l|l|l|l|}
%\hline
% testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
%Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
%Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
%Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
%\end{tabular}
%\end{table}

%\kern-1em
\section{Improvements to TLP for MPTCP}\label{impr}
The goal of Tail Loss Probe~(TLP) is to reduce tail latency of short flows. It achieves this by converting retransmission timeouts (RTOs) occurring due to tail losses into fast recovery. TLP transmits one packet in two round-trips when a connection is in Open state and isn't receiving any ACKs. The transmitted packet, aka loss probe, can be either new or a retransmission. When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout. Our results show that the current TLP implementation for MPTCP helps improving the latency performance in cases when the paths are symmetric or the first path~(for the packet) has lowest RTT. However, it reduces the performance in cases when the first path is not the lowest RTT path or when there is longer path interruption.

The current implementation of TLP is at the TCP flow level. In the event of RTO timeout, the MPTCP retransmission happens with a re-injection in to the scheduler along with sending the lost packet on the same path. But in the event of Probe timeout, the loss probe packet is being sent on the same path without being re-injected in to the scheduler. We propose a modification to the probe sending mechanism for MPTCP by re-injecting the probe packet in to the MPTCP scheduler in the event of PTO. As the probe timeout is meant for acting faster than the RTO timeout with essentially the same consequence as a retransmission, it is relevant to follow the same process as upon an RTO. It is expected that this modification is useful when the alternate path has lower delay than the packet original transmission path or when there is longer path interruption. The timing diagram depicted in Figure~\ref{timingNew} indicates the eventual behavior of modified MPTCP TLP for each of the studied loss scenarios.

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.46\textwidth]{images/timingER3NewTLP1.pdf}
\end{center}
\caption{Timing diagram of MPTCP behavior with proposed improvements with loss of a) Single packet b) Two packets c) Single packet and probe loss}\label{timingNew}
\end{figure}

%\kern-1em
\section{Evaluation of modified MPTCP TLP}\label{eval}

Results comparing the burst completion time for new MPTCP TLP with that of TCP and MPTCP TLP for the considered cases are provided in Figure~\ref{mpnew}. In Figure~\ref{1pn}, MPTCP-New performs better than TCP in all scenarios, performs equal to MPTCP in symmetric scenarios and performs better than MPTCP in asymmetric scenarios with a lower delay on the alternate path. This improvement is for the reason that MPTCP-New allows the packet to be retransmitted on the alternate path much before the RTO. In two packet tail loss scenario as in Figure~\ref{2pn}, MPTCP-New performs better than TCP in all scenarios and better than MPTCP in all scenarios except when the alternate path has a significantly higher delay than the original path, where performance is the same. In the single packet together with probe loss scenario as in Figure~\ref{1ppn}, MPTCP-New performs better than both MPTCP and TCP in all scenarios with PTO triggering the probe retransmission on alternate path much earlier than the RTO.

\begin{figure*}[!tbp]
\begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/1PNew.pdf}
\caption{Single packet}\label{1pn}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/2PNew.pdf}
\caption{Two packets}\label{2pn}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.32\textwidth}
\includegraphics[angle=0, width=\textwidth, natwidth=578.16,natheight=433.62]{plots/1PPNew.pdf}
\caption{Single packet and probe loss}\label{1ppn}
\end{subfigure}
\caption{Performance comparison of MPTCP-New with MPTCP and TCP}\label{mpnew}
\end{figure*}

%\kern-1em
\section{Conclusion}\label{conc}
This paper provides a short comparative analysis of retransmission behavior with TCP and MPTCP protocols using various loss patterns for tail loss.
The three cases of tail loss together with deterministic loss patterns allow us to dissect the retransmission behavior at packet level to understand possible improvements.
We propose a less conservative approach to trigger retransmission on an alternate path in the event of tail loss probe timeout, which otherwise triggered with an RTO. 
Our emulation experiments  using a modified Linux implementation, show that the proposed approach in fact improves the burst completion time in most cases and equals the existing
implementation in other cases. The proposed approach improves the performance when the alternate path has lower delay than the original loss path and the advantage increases with degree of asymmetry between the paths. Temporary path failures can cause probe loss along with packet loss. The approach is very effective in case of probe loss which otherwise incur RTO timeout to trigger retransmission on either path. However, the latency improvement comes at a cost of additional network traffic from the retransmitted packets. Further, we plan to investigate less conservative approaches for other retransmission events.

%\kern-1em
\bibliographystyle{IEEEtran}
\bibliography{mptcp-retrans}
\end{document}